The Adaptive Multi-scale Simulation Infrastructure - Design and Implementation of a soft tissue simulation

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
working on the paper

Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}. Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling equations to take into account the difference in scale, referred to as compression and reconstruction operators in \cite{weinan2003heterogenous} for up-scaling and down-scaling, respectively.

@@

\label{multiscale_paradigms}
Among numerically multi-scale simulations, there are two primary paradigms: concurrent and sequential multi-scale.

The sequential multi-scale paradigm involves the precomputation of parameters using a microscale simulation. These parameters are then used in a constitutive model for the macro-scale simulation \cite{garcia2008sequential}.

@@

@@

The interfaces provided by the amsi libraries are intended to facilitate the coupling of functional (possibly legacy) simulation codes already in usage for single-scale problems. The currently implemented interfaces are physics-agnostic, operating only on computational quantities with no model of a term's relation to the mathematical derivation of the multi-scale interaction. Incorporation of meta-information and services targeted toward specific multi-scale physical couplings represents a desirable extension of the amsi libraries, constituting a rich area of future work.

@@

@@

\label{amsi_communication}
Scale coupling communication is handled by the amsi system using data distributions and communication patterns. A data distribution is a meta representation of individual units of scale-coupling data distributed across a single scale-task. The implementation of the coupling data is arbitrary (so long as it can be serialized for communication). A data distribution represents the smallest unit of data important to the scale-coupling communication, which is necessarily distinct for various multi-scale use cases.

>>>>

Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}. Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling operations to account for differences in dimension as well as possibly dimensionalizing terms computed in a non-dimensional domain.

@@

\label{multiscale_paradigms}
Multi-scale simulations fall broadly into two distinct paradigms concurrent and sequential multi-scale.

@@

In the sequential multi-scale paradigm, tertiary scales compute parameters which are converted to quantities of utility to a primary scale. Each scale must be processed sequentially and is essentially blocked on the completion of the previous scale \cite{garcia2008sequential}. This multi-scale model maps intuitively to phased communication models related to bulk synchronous parallel computation \cite{valiant1990bridging}.

@@

The interfaces provided by the amsi libraries are intended to facilitate the coupling of functional (possibly legacy) simulation codes already in usage for single-scale problems, in order to avoid development overhead associated with casting a simulation into a set of framework-specific components. The currently implemented interfaces are physics-agnostic, operating only on computational quantities with no model of a term's relation to the mathematical derivation of the multi-scale interaction. Incorporation of meta-information and services targeted toward specific multi-scale physical couplings represents a desirable extension of the amsi libraries, constituting a rich area of future work.

@@

\subsection{Novelty} %rename
Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.

These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) in order to take full advantage of such frameworks. Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework shifting some complexity up a level of abstraction. The core features of amsi used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.

simulation flow control is determined solely by the runtime system. A framework may also limit specific features of the numerical implementation to those provided by the framework. Domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc. -- convergence algorithms for both the minimization of the relevant residuals (Newton-Raphson, etc.) as well as iterative solution of the resulting linear system (KSP, GMRES, ILU, etc.). Amsi supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations built on top of other software supplied by SCOREC \cite{core} and the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}. Usage of the supplied numerical components is not required to make use of the multi-scale meta-modeling, communication, dynamic instantiation, and scale-sensitive load balancing facilities supplied by amsi.

While implementation of desired numerical operations into a framework is of course possible, it may prove an additional undesirable overhead for an application programmer interested only in utilizing the multi-scale framework to implement simulation, rather than implementing the various numerical operations often perfected in widely available packages and libraries \cite{} \cite{} \cite{}.

@@

\label{amsi_communication}
Scale coupling communication is guided by data distributions and communication patterns. A data distribution is a representation of individual units of scale-coupling data distributed across a single scale-task. The format of the coupling data is arbitrary (so long as it can be serialized and provided for use by MPI communication primitives). A data distribution represents the smallest unit of data important to the scale-coupling communication, which is necessarily distinct for various multi-scale use cases.
